---
title: "Nupur Nagarkatti – STA 9750 Mini Project"
---
Hi, my name is Nupur Nagarkatti. I am a medical doctor (MBBS) with 5 years of postdoctoral research experience at Yale School of Medicine, where I worked in gynecologic oncology, surgical oncology, and endocrinology.  

My research at Yale sparked a deep interest in applying machine learning and data science to medicine — from analyzing complex datasets like continuous glucose monitoring and neuroimaging, to exploring applications in diagnostics and healthcare innovation.  

I am currently pursuing graduate studies in statistics and data science. This website is part of a course mini-project, and I plan to use it as a space to share my academic journey, projects, and experiments at the intersection of medicine and machine learning.  

```{r}
#| include: false
1+1
```

Last Updated: `r format(Sys.time(), "%A %B %d, %Y at %H:%M %p")`

```{r}
#| echo: false
#| message: false
#| warning: false

if(!require("leaflet")){
    options(repos=c(CRAN="https://cloud.r-project.org"))
    install.packages("leaflet")
    stopifnot(require("leaflet"))
}

baruch_longitude <- -73.98333
baruch_latitude  <- +40.75028

leaflet() |>
  addTiles() |>
  setView(baruch_longitude, baruch_latitude, zoom=17) |>
  addPopups(baruch_longitude, baruch_latitude, 
            "I am a Master's student at <b>Baruch College</b>!")
```


