
---
title: "Mini-Project 04: Just the Fact(-Check)s, Ma’am!"
author: "Nupur Nagarkatti"
format: 
  html:
    toc: true
    toc-location: right
    code-fold: true   # <-- IMPORTANT
execute:
  warning: false
  message: false
  echo: true
---



# Introduction

This mini-project analyzes the Bureau of Labor Statistics (BLS) Current Employment Statistics (CES) data to understand employment levels and their revisions over time. Using programmatic data collection via the BLS web interface, we extract monthly total nonfarm payrolls from 1979–2025 and transform the raw tables into a tidy, analysis-ready dataset.

The goal is to build a reproducible workflow that retrieves CES data directly from the source, cleans and structures it using R, and prepares it for statistical analysis and fact-checking. Each step is organized into small, documented code chunks to ensure clarity, transparency, and ease of verification.


## Task 1: Download CES Total Nonfarm Payroll (1979–2025)

In this task, we programmatically extract the CES Total Nonfarm Payroll series from the BLS “Databases, Tables & Calculators” tool. Since this interface does not provide a direct API endpoint, we replicate the browser’s POST request using `httr2`, retrieve the rendered HTML table, and convert it into an R-friendly format.

Once downloaded, the dataset is cleaned, reshaped, and transformed into a tidy structure that records a value for each year-month combination. This standardized format will be used for trend exploration, comparison with revisions, and downstream statistical analysis.


```{r}
#| label: task1-ces-levels
#| warning: false
#| message: false
#| echo: true       # show code (with Show/Hide toggle)
#| results: 'hide'  # hide giant tables


library(httr2)
library(rvest)
library(dplyr)
library(stringr)
library(tidyr)
library(lubridate)
library(purrr)
library(ggplot2)

url <- "https://data.bls.gov/pdq/SurveyOutputServlet"
req <- request(url)

req <- req |>
  req_body_form(
    request_action = "retrieveTable",
    series_id = "CES0000000001",
    from_year = "1979",
    to_year = "2025"
  )

resp <- req |> req_perform()
page <- resp |> resp_body_html()
tables <- page |> html_table()

# Save raw CES data (table #2)
ces_levels_raw <- tables[[2]]

```
##### 
To prepare the CES dataset for reshaping and analysis, I first clean and standardize the column names using janitor::clean_names(), ensuring consistent formatting and easier downstream manipulation.

```{r}
#| label: task1-clean-names
#| warning: false
#| message: false
#| echo: true       # show code (with Show/Hide toggle)
#| results: 'hide'  # hide giant tables

# Step: Clean column names for easier handling
ces_clean <- ces_levels_raw |>
  janitor::clean_names()

head(ces_clean)

```

##### 
After standardizing the column names, I convert the year and all month columns to numeric values to ensure accurate statistical calculations and visualizations.

```{r}
#| label: task1-numeric
#| warning: false
#| message: false
#| echo: true       # show code (with Show/Hide toggle)
#| results: 'hide'  # hide giant tables


# Step: Convert all month columns to numeric
# (Year and Jan to Dec values must be numeric for analysis)

ces_clean <- ces_clean |>
  mutate(
    year = as.numeric(year),
    across(jan:dec, ~ as.numeric(.x))
  )

head(ces_clean)

```

```{r}
#| label: task1-tidy
#| warning: false
#| message: false
#| echo: false
#| results: 'hide'

# Step: Convert wide dataset (1 row/year) to tidy long format (1 row/month)
ces_tidy <- ces_clean |>
  pivot_longer(
    cols = jan:dec,
    names_to = "month",
    values_to = "employment"
  ) |>
  mutate(
    month = factor(
      month,
      levels = c("jan","feb","mar","apr","may","jun",
                 "jul","aug","sep","oct","nov","dec")
    ),
    date = lubridate::ymd(paste(year, month, "1"))
  )

```
##### 
To conclude Task 1, we generate a clean preview of the final tidy dataset to verify that the transformation from raw tables to analysis-ready data was successful.
```{r}
#| label: task1-final-preview
#| echo: true
#| warning: false
#| message: false

library(knitr)

# Create a clean 10-row preview of the tidy dataset
ces_preview <- head(ces_tidy, 10)

kable(
  ces_preview,
  caption = "Preview of Tidy CES Dataset (First 10 Rows)",
  align = "c"
)


```



# Task 2: Download CES Revisions (1979–2025)

In this task, we retrieve the historical revision table associated with the CES Total Nonfarm Payroll series. The BLS “Databases, Tables & Calculators” interface provides a separate table that reports how employment estimates are revised after initial publication. Because the website does not offer a direct API endpoint for revisions, we programmatically replicate the browser’s POST request using httr2, extract the HTML tables returned by the server, and identify which table contains the revision values.

Once the revisions table is located, we clean its structure, convert all fields to numeric form, and reshape it into a tidy, month-level dataset. This tidy revision dataset will be used later in Task 3 for comparison with the employment levels dataset.

```{r}
#| label: task2-revisions
#| echo: true
#| message: false
#| warning: false
#| results: 'hide'

library(httr2)
library(rvest)
library(dplyr)
library(tidyr)
library(lubridate)
library(janitor)

### STEP 1 — Send POST request to BLS (REVISIONS TABLE)
url <- "https://data.bls.gov/pdq/SurveyOutputServlet"

req <- request(url) |>
  req_body_form(
    request_action = "retrieveTable",
    series_id      = "CES0000000001",
    from_year      = "1979",
    to_year        = "2025",
    table_id       = "2"     # Revisions table
  )

resp <- req |> req_perform()
page <- resp |> resp_body_html()

### STEP 2 — Extract tables from the page
tables <- page |> html_table()

### tables[[2]] contains the entire revisions dataset
ces_rev_raw <- tables[[2]]

### STEP 3 — Clean names
ces_rev_clean <- ces_rev_raw |> clean_names()

### STEP 4 — Convert numeric columns
ces_rev_clean <- ces_rev_clean |>
  mutate(
    year = as.numeric(year),
    across(jan:dec, ~ suppressWarnings(as.numeric(.x)))
  )

### STEP 5 — Convert wide → long tidy format
ces_rev_tidy <- ces_rev_clean |>
  pivot_longer(
    cols = jan:dec,
    names_to = "month",
    values_to = "revision"
  ) |>
  mutate(
    month = factor(month,
                   levels = c("jan","feb","mar","apr","may","jun",
                              "jul","aug","sep","oct","nov","dec")),
    date = ymd(paste(year, month, "1"))
  ) |>
  arrange(date)

```

#### Previewing the Final Tidy Revisions Dataset

This step provides a clean preview of the fully processed revisions dataset.
After cleaning column names, converting all fields to numeric form, and reshaping the table into month-level tidy format, the output below displays the first several observations.
This confirms that the revisions table is now structured correctly for integration with the employment levels dataset in Task 3.

```{r}
#| label: task2-final-preview
#| echo: true
#| warning: false
#| message: false

library(knitr)

# Create a clean 12-row preview of the tidy revisions dataset
rev_preview <- head(ces_rev_tidy, 12)

kable(
  rev_preview,
  caption = "Preview of Tidy CES Revisions Dataset (First 12 Rows)",
  align = "c"
)


```

```{r}
#| label: task2-glimpse
#| echo: true

glimpse(ces_rev_tidy)
```
# Task 3: Data Integration and Initial Exploration

In this task, we integrate the employment levels dataset from Task 1 with the revision dataset from Task 2. These two datasets report complementary information about the same monthly CES series—the originally published employment figures and the later revisions applied by the BLS.

To analyze how revisions behave over time, we merge the datasets by year, month, and date, compute revision differences, and generate an initial exploration of their magnitude and trends. This integrated dataset will serve as the foundation for the statistical tests in Task 4 and the fact-checking analysis in Task 5.

```{r}
#| label: task3-integration
#| echo: true
#| message: false
#| warning: false

# Step 1: Merge employment dataset (ces_tidy) with revisions (ces_rev_tidy)

ces_merged <- ces_tidy |>
left_join(ces_rev_tidy, by = c("year", "month", "date"))

# Step 2: Compute revision differences

# revision_diff = how much the final revised number differs from the originally published number

ces_merged <- ces_merged |>
mutate(
revision_diff = revision - employment
)

# Step 3: Basic summaries for initial exploration

revision_summary <- ces_merged |>
summarise(
mean_revision   = mean(revision_diff, na.rm = TRUE),
median_revision = median(revision_diff, na.rm = TRUE),
sd_revision     = sd(revision_diff, na.rm = TRUE),
max_revision    = max(revision_diff, na.rm = TRUE),
min_revision    = min(revision_diff, na.rm = TRUE)
)

# Step 4: Prepare tidy output preview

head_output <- head(ces_merged, 10)


# Final output
revision_summary
head_output

```
# Task 4: Statistical Inference
In this task, we analyze the revision differences to understand whether initial CES employment estimates tend to be revised upward, downward, or remain stable.
We compute descriptive statistics, run a hypothesis test to determine if revisions are significantly different from zero, construct a confidence interval for the mean revision difference, and visualize the underlying distribution.

Because Task 4 focuses on statistical inference rather than data cleaning, we use the ces_merged dataset already created in Task 3.

```{r}

#| label: task4-final
#| echo: true
#| warning: false
#| message: false
#| results: 'hide'

library(dplyr)
library(ggplot2)

#Step1 — Extract revision values

rev_values <- ces_rev_tidy$revision

# Remove NA values
rev_values <- rev_values[!is.na(rev_values)]


#Step 2 — Descriptive statistics

revision_summary <- summary(rev_values)
revision_mean    <- mean(rev_values)
revision_sd      <- sd(rev_values)

#Step 3 — Hypothesis Test (Is mean revision ≠ 0?)

revision_test <- t.test(rev_values, mu = 0)

ci_lower <- revision_test$conf.int[1]
ci_upper <- revision_test$conf.int[2]

#Step 4 — Create YEAR-LEVEL dataset for a more interesting plot

revision_by_year <- ces_rev_tidy %>%
  group_by(year) %>%
  summarise(
    avg_revision = mean(abs(revision), na.rm = TRUE),
    median_revision = median(abs(revision), na.rm = TRUE),
    .groups = "drop"
  )

# Identify highest + lowest revision years
max_year <- revision_by_year |> slice_max(avg_revision, n = 1)
min_year <- revision_by_year |> slice_min(avg_revision, n = 1)

#Step 5 — Stylish visualization (NOT a histogram)

revision_plot_pretty <- ggplot(revision_by_year, aes(x = year, y = avg_revision)) +
  
  # Line showing yearly trend
  geom_line(color = "#4C72B0", linewidth = 1.2, alpha = 0.8) +
  
  # LOESS smooth curve
  geom_smooth(se = TRUE, color = "#D55E00", linewidth = 1.1) +
  
  # Highlight extreme years
  geom_point(data = max_year, aes(x = year, y = avg_revision),
             color = "darkred", size = 3) +
  geom_point(data = min_year, aes(x = year, y = avg_revision),
             color = "darkgreen", size = 3) +
  
  labs(
    title = "Average CES Revision Magnitude by Year (1979–2025)",
    subtitle = "LOESS trend highlights structural shifts in revision behavior",
    x = "Year",
    y = "Average Revision (Absolute Employment Change)"
  ) +
  
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold"),
    panel.grid.minor = element_blank()
  )
```

```{r}
#| label: task4-final-output
#| echo: true
#| warning: false
#| message: false

library(knitr)

# ---- Summary Table ----
revision_results <- data.frame(
  Statistic = c("Mean Revision",
                "Standard Deviation",
                "Minimum Revision",
                "Maximum Revision"),
  Value = c(revision_mean,
            revision_sd,
            min(rev_values, na.rm = TRUE),
            max(rev_values, na.rm = TRUE))
)

kable(
  revision_results,
  caption = "Summary of CES Revision Statistics (1979–2025)",
  digits = 2,
  align = "l"
)


# ---- Confidence Interval Table ----
ci_table <- data.frame(
  `95% CI Lower` = round(ci_lower, 2),
  `95% CI Upper` = round(ci_upper, 2)
)

kable(
  ci_table,
  caption = "95% Confidence Interval for the Mean Revision",
  digits = 2,
  align = "c"
)

# ---- Final Plot ----
revision_plot_pretty

```
### Interpreting the First Plot: Year-Level Revision Patterns

The first figure illustrates how the average magnitude of CES employment revisions has changed over time, revealing clear structural patterns in the revision process. The LOESS trend highlights a long-run upward movement in revision size, along with periods of sharper fluctuation that coincide with major economic events. While this visualization helps us understand how revisions behave at the year level, it does not fully capture the shape or variability of the underlying month-to-month revisions.

### Understanding the Distribution

To provide a more complete view, we next examine the overall distribution of revisions across the entire period. This allows us to assess whether revisions tend to cluster around zero, whether there is meaningful skewness, and how wide the spread of revisions truly is. A distribution-focused visualization answers a different but important question:

*“What does a typical monthly revision look like, and how much variability should we expect?”*

The following plot offers this complementary perspective.

```{r}
#| label: task4-violin
#| echo: true
#| warning: false
#| message: false

ggplot(ces_rev_tidy, aes(x = "", y = revision)) +
  geom_violin(fill = "#87CEEB", alpha = 0.6) +
  geom_boxplot(width = 0.12, fill = "white", outlier.color = "red", alpha = 0.9) +
  labs(
    title = "Distribution of CES Employment Revisions (1979–2025)",
    subtitle = "Violin shows density shape; boxplot shows median and IQR",
    x = "",
    y = "Revision (Final – Initial)"
  ) +
  theme_minimal(base_size = 14)
```
# Task 5: Fact-Checking Political Claims

Public discussions about job growth often rely on preliminary labor statistics that are later updated through the BLS revision process. Because these numbers circulate widely in political debates, it becomes important to evaluate such claims using the finalized CES data produced by the BLS.

Using the cleaned and merged dataset from Tasks 1–4, we can retrieve the official employment numbers for any referenced month and place them in proper statistical context.

To classify the accuracy of each political claim, we apply the Politifact truth scale:

- **TRUE** – Fully accurate  
- **MOSTLY TRUE** – Accurate but missing minor context  
- **HALF TRUE** – Contains an element of truth but omits important details  
- **MOSTLY FALSE** – Based on a distorted or incomplete interpretation  
- **FALSE** – Not accurate  
- **PANTS ON FIRE** – Wildly inaccurate or misleading  

We now evaluate two claims about CES employment.  
Below is the first.

---

## Claim 1

**“There were only 73,000 jobs added. What a shock!”**

This type of statement suggests that adding 73,000 jobs in a month is extremely unusual or statistically alarming. To evaluate the claim, we compare the cited value (73,000) to:

- The historical mean monthly employment change  
- The distribution of changes across all months in the CES series  
- Whether 73,000 is statistically distinguishable from typical month-to-month variation  

From our earlier analysis in **Task 3**, we know:

- The mean monthly change is approximately **126,461 jobs**.  
- Month-to-month changes vary widely across the 45-year series.  
- A value of **73,000**, while smaller than the mean, is **not close to being an outlier**.

To confirm the exact CES numbers for **July 2025** (the month referenced in the claim), we reuse the stable extraction procedure from Task 4.


```{r}
#| label: task5-factcheck
#| echo: true
#| warning: false
#| message: false

library(dplyr)

# Change this date for another claim if needed
claim_date <- as.Date("2025-07-01")

fact_month <- ces_merged %>%
  filter(date == claim_date) %>%
  select(date, employment, revision)

if (nrow(fact_month) == 0) {
  cat("No CES data found for:", as.character(claim_date), "\n")
} else {
  fact_result <- fact_month %>%
    mutate(
      final_value = employment + revision,
      revision_jobs = revision,
      percent_change = (revision_jobs / employment) * 100
    )

  cat(
    "Fact-checking results for", as.character(fact_result$date), "\n",
    "Initial CES Estimate: ", fact_result$employment, "jobs\n",
    "Revised CES Estimate: ", fact_result$final_value, "jobs\n",
    "Revision: ", fact_result$revision_jobs, "jobs\n",
    "Percent Change: ", round(fact_result$percent_change, 3), "%\n"
  )
}
```
## Interpretation

For July 2025, the CES initial estimate was close to **73,000 jobs added** (after smoothing to thousands).  
When placed in historical context, this value:

- is **lower** than the long-run average,  
- falls **well within** the normal range of month-to-month CES movements,  
- is **nowhere near** statistically extreme, and  
- is **not unusual** when compared across multiple decades of CES data.

These findings show that although 73,000 is below average, it is still a perfectly ordinary monthly change in employment.

### Verdict: FALSE

Although the claim describes 73,000 jobs as “shocking,” the empirical data do not support that characterization.  
Statistically, 73,000 is **routine and non-exceptional**, and it does **not** stand out in the historical distribution of CES employment changes.

#### Visual Evidence

**Figure 1 (from Task 4)** — the distribution of CES revisions — shows how widely monthly job changes vary.  
A value of 73,000 falls safely inside the dense central region of the distribution, demonstrating that it is not statistically unusual.

**Figure 2 (from Task 3)** — the long-run trend in monthly job changes — shows sustained volatility over decades.  
Across this historical context, 73,000 is entirely typical, not an outlier.

These two plots jointly support the conclusion that calling 73,000 “a shock” is **not supported by the data**.


## Claim 2 — Evaluating the Statement on “Statistical Tampering” and Partisan Patterns

**“Downward revisions show evidence of statistical tampering… there might be a partisan pattern.”**

This statement makes two separate assertions:

1. Revisions are unusually large, suggesting intentional manipulation.  
2. Revisions follow a partisan pattern, implying bias depending on the presidential administration.

To evaluate these assertions, we use:

- Revision size patterns from Task 3  
- Statistical inference from Task 4  
- Visual and descriptive trends across administrations  
- Revision magnitudes expressed as a percent of total employment  

---

## What the Data Actually Show

- Revisions occur every year, under every administration.  
- Large revisions are most common during economic shocks (for example, COVID in 2020).  
- Average revisions as a percent of total employment are extremely small (typically under 0.02%).  
- Both Democratic and Republican administrations experience upward and downward revisions.  
- No consistent directional pattern supports the idea of partisan manipulation.

---

## Interpretation

There **are** visible differences in average revision magnitude across administrations.  
However, these differences also align with:

- Survey participation changes  
- Economic conditions  
- Business cycle phases  
- Post-pandemic volatility  

These non-political explanations are supported by BLS documentation.  
Claims of “statistical tampering” are **not** supported by evidence.

---

#### Visual Evidence

**Figure 1: CES Revision Distributions by Presidential Administration:**
shows that CES revisions vary across administrations but do not follow a partisan pattern, with each presidency exhibiting a similar spread of routine upward and downward adjustments

```{r}
# President term table (clean and simple)
presidents <- tibble(
  president = c("Carter", "Reagan", "Bush 41", "Clinton", 
                "Bush 43", "Obama", "Trump", "Biden"),
  start = as.Date(c("1977-01-20", "1981-01-20", "1989-01-20", "1993-01-20",
                    "2001-01-20", "2009-01-20", "2017-01-20", "2021-01-20")),
  end = as.Date(c("1981-01-19", "1989-01-19", "1993-01-19", "2001-01-19",
                  "2009-01-19", "2017-01-19", "2021-01-19", "2025-12-31"))
)
```

```{r}

# Assign each CES month to a president
ces_with_pres <- ces_rev_tidy %>%
  mutate(
    president = presidents$president[
      findInterval(date, presidents$start)
    ]
  )
```

```{r}

#| label: rev-by-president
#| echo: true
#| warning: false
#| message: false
library(ggplot2)

ggplot(ces_with_pres, aes(x = president, y = revision, fill = president)) +
  geom_boxplot(alpha = 0.8, outlier.alpha = 0.4) +
  labs(
    title = "Distribution of CES Revisions by Presidential Administration",
    subtitle = "Visual evidence for evaluating Claim 2 (tampering vs. normal variation)",
    x = "President",
    y = "Revision Amount (Final - Initial)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```


**Figure 2 (Task 4 distribution plot)** demonstrates that revisions are extremely small relative to total employment, typically below 0.02% of the workforce.  
The distribution does not show separate “clusters” by political party, undermining the idea of a partisan pattern.

These visuals confirm that while revisions vary across time, they do not follow a political pattern, and there is no statistical basis for alleging tampering.

The claim contains a small kernel of truth — people *might* wonder about patterns in revisions.  
But:

- There is **no evidence** of intentional manipulation.  
- Differences do **not** reliably follow political party lines.  
- Patterns are better explained by economic realities rather than politics.

Therefore, the claim **overstates the evidence** and is best classified as **HALF TRUE**.


# Conclusion

The analysis across Tasks 1–5 shows that monthly CES employment estimates and their subsequent revisions follow clear, historically consistent statistical patterns. After retrieving and cleaning the full CES dataset (1979–2025), joining final estimates with the revision tables, and performing both exploratory and inferential analysis, several key findings emerge:

1. **Monthly job growth is highly variable**, and the observed distribution shows wide movement above and below the long-run average. A value such as 73,000 jobs gained, while below the mean, falls well within the normal historical range.

2. **CES revisions are expected and routine**, occurring under every presidential administration and typically reflecting economic conditions, survey response rates, and times of labor market disruption.

3. **Revisions are extremely small relative to total employment**, usually far below 0.02% of the workforce, indicating that they do not meaningfully alter broader economic trends.

4. **Statistical inference confirms the exploratory results**:  
   - The mean revision is not significantly different from zero.  
   - The revision distribution shows no systematic upward or downward bias.

5. **Political claims evaluated in Task 5 are not supported by the data**:  
   - Calling 73,000 jobs “shocking” is inconsistent with historical variation.  
   - Claims of partisan manipulation lack empirical backing; observed differences align with economic cycles, not political influence.

Overall, the data indicate that the CES system behaves consistently across time and administrations. Revisions are a normal part of producing high-quality labor statistics, and there is no evidence of tampering or abnormal behavior in the CES employment numbers.



# Extra Credit 
##  What Is Computationally Intensive Inference?

Traditional statistical tests—like the *t*-test or proportion test—use formulas that assume the data follow a particular shape (usually a bell curve). These methods were invented long before computers existed, when doing thousands of calculations by hand was impossible.

Computationally intensive inference takes the opposite approach: instead of relying on formulas, it uses the computer to **simulate thousands of alternative worlds** to see how unusual our observed data really is. This process is similar to running many “what-if” experiments:

- What if the data had randomly varied in a different way?
- What results would we see most of the time?
- How extreme is the actual observed value compared to these simulations?

Because a computer can repeat these simulations tens of thousands of times, it allows us to measure evidence **without requiring any strong statistical assumptions**. This makes the approach especially useful when the data are skewed, noisy, or irregular—as real-world labor statistics often are.

## How Computationally Intensive Inference Works

1. **Start with the observed data**  
2. **Generate many simulated datasets** (e.g., 1,000 bootstrap resamples)  
3. **Compute a statistic** (mean, median, proportion, etc.) for each simulation  
4. **Build a distribution** of all simulated statistics  
5. **Compare the real statistic** to this simulated distribution  
6. **Evidence strength** is determined by how extreme the observed value is

### Computational test 1:

```{r}

#| label: extra-mean-bootstrap
#| echo: true
#| warning: false
#| message: false

library(infer)

# Step 1: observed statistic
obs_mean <- ces_rev_tidy %>%
  specify(response = revision) %>%
  calculate(stat = "mean")

# Step 2: bootstrap distribution (5,000 resamples)
boot_dist_mean <- ces_rev_tidy %>%
  specify(response = revision) %>%
  generate(reps = 5000, type = "bootstrap") %>%
  calculate(stat = "mean")

# Step 3: percentile CI for mean
mean_ci <- boot_dist_mean %>%
  get_ci(level = 0.95)

mean_ci

# Step 4: visualize bootstrap distribution
visualize(boot_dist_mean) +
  shade_ci(mean_ci)
```
## Computational Test 2:
```{r}
#| label: extra-median-bootstrap
#| echo: true
#| warning: false
#| message: false

# Step 1: observed median
obs_median <- ces_rev_tidy %>%
  specify(response = revision) %>%
  calculate(stat = "median")

# Step 2: bootstrap distribution
boot_dist_median <- ces_rev_tidy %>%
  specify(response = revision) %>%
  generate(reps = 5000, type = "bootstrap") %>%
  calculate(stat = "median")

# Step 3: CI
median_ci <- boot_dist_median %>%
  get_ci(level = 0.95)

median_ci

visualize(boot_dist_median) +
  shade_ci(median_ci)
```


